---
title: "TMA syntax design: accumulation using context impact function"
output: html_notebook
---
```{r}
df <- read.csv(file = "mock_data_edited.csv")
```

Old syntax
```{r}
unit_cols <- c("userID")

codes <- c("A", "B", "C")

HOO_rules_model <- tma:::rules(
  modality %in% "chat" & chatGroup %in% UNIT$chatGroup & Condition %in% UNIT$Condition, 
  modality %in% "resource" & userID %in% UNIT$userID & Condition %in% UNIT$Condition
)

context_model <- tma:::contexts(
  x = df,
  units = unit_cols,
  hoo_rules = HOO_rules_model
)

accum_unordered <- tma:::accumulate_contexts (
    x = context_model,
    codes = codes,
    weight.by = function(x) { x },
    decay.function = function(x) {
        modality_effect <- c("chat" = 2, "resource" = 4) 
        modality <- FULL_CONTEXT[ROWS, modality]
        weight <- (x < (modality_effect[modality])) * 1
        return (weight)
    },
    time.column = "timeStamp",
    return.ena.set = TRUE
)

```

Solution 1: Multidimensional array based 

Shortcut of TIF only (modify window length based on modalities)
```{r}
unit_cols <- c("userID")

codes <- c("A", "B", "C")

HOO_rules_model <- tma:::rules(
  modality %in% "chat" & chatGroup %in% UNIT$chatGroup & Condition %in% UNIT$Condition, 
  modality %in% "resource" & userID %in% UNIT$userID & Condition %in% UNIT$Condition
)

context_model <- tma:::contexts(
  x = df,
  units = unit_cols,
  hoo_rules = HOO_rules_model
)

accum_unordered <- tma:::accumulate_contexts (
    x = context_model,
    codes = codes,
    time.column = "timeStamp",
    return.ena.set = TRUE,
    based.window = 0,
    mode.column = "modality",
    mode.values = c("chat" = 2, "resource" = 4)
)
```

Most general case:
```{r}
unit_cols <- c("userID")

codes <- c("A", "B", "C")

RS.hoo.cif.combined.rules = list(
  cif.base.window = 4, 
  cif.base.weight = 1,
  cif.mode.column = "modality", 
  cif.sender.column = c("role", "language", "handwriting", "userID", "chatGroup"), 
  cif.receiver.column = c("language", "handwriting", "userID", "chatGroup"), 
  cif.exception.rule = tma:::cif_exceptions(
    rule_1 = list(adjust.type = window.type, 
                  adjust.num = 1,
                  adjust.element = c("resource", "Player", "L2","poor", NA, NA, 
                                     "L1", "good", NA, NA)
                  ),
    rule_2 = list(adjust.type = window.type, 
                  adjust.num = 2,
                  adjust.element = c("chatting",NA,NA,NA, NA,NA, 
                                     "L2",NA,NA,NA)
                  ),
    rule_3 = list(adjust.type = weight.type, 
                  adjust.num = 0.5,
                  adjust.element = c("chatting", "Player", "L2", "poor",NA,NA,
                                     NA,NA,NA,NA)
                  ), 
    rule_4 = list(adjust.type = weight.type, 
                  adjust.num = 2,
                  adjust.element = c("chatting", "Mentor",NA, "good", NA,NA,
                                     "L2",NA, NA,NA)
                  ), 

    rule_5 = list(adjust.type = weight.type, 
                  adjust.num = 0,
                  adjust.element = c(NA, NA,NA, NA,NA,chatGroup,
                                     NA,NA, NA,chatGroup), 
                  unmatched.element = c(NA,NA,NA,NA,NA,TRUE,
                                      NA,NA,NA,TRUE)
                  ), 
    rule_6 = list(adjust.type = weight.type, 
                  adjust.num = 0,
                  adjust.element = c("resource", NA,NA, NA,userID,chatGroup,
                                     NA,NA, userID,chatGroup), 
                  matched.element = c(NA,NA,NA,NA,NA,TRUE,
                                      NA,NA,NA,TRUE), 
                  unmatched.element = c(NA,NA,NA,NA,TRUE,NA,
                                      NA,NA,TRUE,NA)
                  )
    )
)
accum_unordered <- tma:::accumulate_contexts (
    x = df,
    codes = codes,
    time.column = "timeStamp",
    return.ena.set = TRUE,
    hoo.rules = RS.hoo.rules, 
    cif.rules = RS.cif.rules
)
```
```{r}
unit_cols <- c("userID")

codes <- c("A", "B", "C")

HOO_rules_model <- tma:::rules(
  modality %in% "chat" & chatGroup %in% UNIT$chatGroup & Condition %in% UNIT$Condition, 
  modality %in% "resource" & userID %in% UNIT$userID & Condition %in% UNIT$Condition
)

context_model <- tma:::contexts(
  x = df,
  units = unit_cols,
  hoo_rules = HOO_rules_model
)

accum_unordered <- tma:::accumulate_contexts (
    x = context_model,
    codes = codes,
    time.column = "timeStamp",
    return.ena.set = TRUE,
    base.window = 4, 
    base.weight = 1,
    # adjust.type = c(window.type,weight.type), # adjust.type = window.type or adjust.type = weight.type 
    cif.mode.column = "modality", 
    cif.sender.column = c("role", "language", "handwriting"), 
    cif.receiver.column = c("language", "handwriting"), 
    cif.exception.rule = tma:::cif_exceptions(
      rule_1 = list(adjust.type = window.type, 
                    adjust.num = 1,
                    adjust.element = c("resource", "Player", "L2", "poor", "L1", "good")
                   ),
      rule_2 = list(adjust.type = window.type, 
                    adjust.num = 2,
                    adjust.element = c("chatting",NA,NA,NA,"L2",NA)
                    ),
      rule_3 = list(adjust.type = weight.type, 
                    adjust.num = 0.5,
                    adjust.element = c("chatting", "Player", "L2", "poor",NA,NA), 
                    ), 
      rule_4 = list(adjust.type = weight.type, 
                    adjust.num = 2,
                    adjust.element = c("chatting", "Mentor",NA, "good", "L2",NA), 
                    )
      )
)
```
Based on the variables specified for mode.column, sender.column, and receiver.column, TMA creates a multidimensional array to record windows and weights for different contexts. The dimension $d$ of this multidimensional array is the same as total number of variables specified for the three arguments. 

Error scenarios:
1. If the length of adjust.element vectors in exception.rule is not matched with $d$, throw error message "Incorrect dimension specification."
2. If two exception rules try to modify the same element in the multidimensional array to two different values, throw error message "Conflicts in modifying exceptions."
3. If the values adjust.element vectors in the exception.rule don't match with the unique values of the corresponding specified variable, throw error message "Value not found in the specified column for mode.column/sender.column/receiver.column."


Solution 2: Rule based

```{r}
unit_cols <- c("userID")

codes <- c("A", "B", "C")

RS.hoo.cif.combined.rules = list(
  cif.base.window = 4, 
  cif.base.weight = 1,
  cif.mode.column = "modality", 
  cif.sender.column = c("role", "language", "handwriting", "userID", "chatGroup"), 
  cif.receiver.column = c("language", "handwriting", "userID", "chatGroup"), 
  cif.exception.rule = tma:::cif_exceptions(
    if(modality == "resource" & sender.role == "Player" & sender.language == "L2" & 
       sender.handwriting == "poor" & receiver.language == "L1" & receiver.handwriting == "good") {
      adjust.window = 1
    }, 
    if (modality == "chatting" & receiver.language == "L2") {
      adjust.window = 2
    },
    if (modality == "chatting" & sender.role == "Player" & sender.language == "L2" & sender.handwriting == "poor") {
      adjust.weight = 0.5
    },
    if (modality == "chatting" & sender.role == "Mentor" & sender.handwriting == "good" & receiver.language == "L2") {
      adjust.weight = 2
    },
    if (modality == "chat" & sender.chatGroup != receiver.chatGroup) {
      adjust.weight = 0
    },
    if (modality == "resource" & (sender.userID != receiver.userID | sender.chatGroup == receiver.chatGroup))     {
      adjust.weight = 0
    }
  )
)
accum_unordered <- tma:::accumulate_contexts (
    x = df,
    codes = codes,
    time.column = "timeStamp",
    return.ena.set = TRUE,
    hoo.rules = RS.hoo.rules, 
    cif.rules = RS.cif.rules
)
```
